A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-04-09 paper.Rmd
slight word-smitting of intro

We can assess residuals in multiple ways. Residuals may be plotted as a histogram or quantile-quantile plot to examine the distribution. Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we consider it to be well behaved. However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there is a cause for concern. One could also inspect the distribution by conducting a goodness of fit test, such as the Shapiro-Wilk (SW) Normality test [@shapiro1965analysis].

More typically, residuals will be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships. If there are any visually discoverable patterns, the model is potentially misspecified. In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity. However, correctly judging a residual plot where no pattern exists can be a painstakingly difficult task for humans. It is especially common, particularly among students, to misinterpret patterns that are random noise and random deviation from a model [@loy2021bringing]. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan (BP) test [@breusch_simple_1979] for heteroskedasticity.

Abundance of literature describe appropriate diagnostic methods for linear regression, e.g. @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals. All these writings consider plotting residuals to be a standard technique that should be examined routinely in all regression modelling problems. In addition, @draper1998applied and @belsley_regression_1980 believe that residual plots are usually revealing when the assumptions are violated. @cook_applied_1999 thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing, as they are easier to use. @montgomery1982introduction even suggests that residual plots are more informative in most practical situations than the corresponding formal tests, and statistical tests on regression model residuals are not widely used based on their experience.

@@

The power of a visual test on the other hand, may additionally depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013]. To address this issue, @majumder_validation_2013 models the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effect. Then, the estimated power of a visual test evaluated by a single subject is the predicted value obtained from the mix-effect model. However, this mix-effect model does not work with scenario where subjects are asked to select one or more most different plots. In this scenario, having the probability of a subject $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the $p$-value (See Equation \ref{eq:pvalue-beta-binomial-approx}). Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power. This assumption is not necessary true, but it helps simplifying the model structure, thereby obviate a costly large-scale experiment to estimate complex covariance matrices. The same model given in Equation \ref{eq:logistic-regression-1-1} is applied to model the power of a visual test.

>>>>

We can assess residuals in multiple ways. Residuals may be plotted as a histogram or quantile-quantile plot to examine the distribution. Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we consider it to be well behaved. However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there is a cause for concern. One could also inspect the distribution by conducting a goodness-of-fit test, such as the Shapiro-Wilk (SW) Normality test [@shapiro1965analysis].

More typically, a scatterplot of residuals plotted against the fitted values and each of the explanatory variables, is made to scrutinize their relationships. If there are any visually discoverable patterns, the model is potentially inadequate or incorrectly specifed. In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity. However, correctly judging a residual plot where no pattern exists can be a difficult task for humans. It is especially common, particularly among students, to misinterpret patterns that are random noise and random deviation from a model [@loy2021bringing]. There are also numerous conventional hypothesis tests available to test for non-linear dependence [@ramsey_tests_1969], and heteroskedasticity  [@breusch_simple_1979].

An abundance of literature describe appropriate diagnostic methods for linear regression, e.g. @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals. All these writings consider plotting residuals to be an essential tool for diagnosing regression modelling problems. In addition, @draper1998applied and @belsley_regression_1980 believe that residual plots are usually revealing when the assumptions are violated. @cook_applied_1999 thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing, as they are easier to use. @montgomery1982introduction even suggests that residual plots are more informative in most practical situations than the corresponding conventional hypothesis tests.

@@

The power of a visual test on the other hand, may additionally depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013]. To address this issue, @majumder_validation_2013 models the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effect. Then, the estimated power of a visual test evaluated by a single subject is the predicted value obtained from the mixed effects model. However, this mixed effects model does not work with scenario where subjects are asked to select one or more most different plots. In this scenario, having the probability of a subject $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the $p$-value (See Equation \ref{eq:pvalue-beta-binomial-approx}). Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power. This assumption is not necessary true, but it helps simplifying the model structure, thereby obviate a costly large-scale experiment to estimate complex covariance matrices. The same model given in Equation \ref{eq:logistic-regression-1-1} is applied to model the power of a visual test.

### 2023-04-09 paper.tex
slight word-smitting of intro

We can assess residuals in multiple ways. Residuals may be plotted as a
histogram or quantile-quantile plot to examine the distribution. Using
the classical normal linear regression model as an example, if the
distribution is symmetric and unimodal, we consider it to be well
behaved. However, if the distribution is skewed, bimodal, multimodal, or
contains outliers, there is a cause for concern. One could also inspect
the distribution by conducting a goodness of fit test, such as the
Shapiro-Wilk (SW) Normality test \citep{shapiro1965analysis}.

More typically, residuals will be plotted, as a scatter plot against the
predicted values and each of the explanatory variables to scrutinize
their relationships. If there are any visually discoverable patterns,
the model is potentially misspecified. In general, one looks for
noticeable departures from the model like non-linear dependency or
heteroskedasticity. However, correctly judging a residual plot where no
pattern exists can be a painstakingly difficult task for humans. It is
especially common, particularly among students, to misinterpret patterns
that are random noise and random deviation from a model
\citep{loy2021bringing}. It is also possible to conduct hypothesis tests
for non-linear dependence \citep{ramsey_tests_1969}, and use a
Breusch-Pagan (BP) test \citep{breusch_simple_1979} for
heteroskedasticity.

Abundance of literature describe appropriate diagnostic methods for
linear regression, e.g. \citet{draper1998applied},
\citet{montgomery1982introduction}, \citet{belsley_regression_1980},
\citet{cook_applied_1999} and \citet{cook1982residuals}. All these
writings consider plotting residuals to be a standard technique that
should be examined routinely in all regression modelling problems. In
addition, \citet{draper1998applied} and \citet{belsley_regression_1980}
believe that residual plots are usually revealing when the assumptions
are violated. \citet{cook_applied_1999} thinks formal tests and
graphical procedures are complementary and both have a place in residual
analysis, but they focus on graphical methods rather than on formal
testing, as they are easier to use. \citet{montgomery1982introduction}
even suggests that residual plots are more informative in most practical
situations than the corresponding formal tests, and statistical tests on
regression model residuals are not widely used based on their
experience.

@@

The power of a visual test on the other hand, may additionally depend on
the ability of the particular subject, as the skill of the individual
may affect the number of observers who identify the data plot from the
lineup \citep{majumder_validation_2013}. To address this issue,
\citet{majumder_validation_2013} models the probability of a subject
\(j\) correctly picking the data plot from a lineup \(l\) using a
mixed-effect logistic regression, with subjects treated as random
effect. Then, the estimated power of a visual test evaluated by a single
subject is the predicted value obtained from the mix-effect model.
However, this mix-effect model does not work with scenario where
subjects are asked to select one or more most different plots. In this
scenario, having the probability of a subject \(j\) correctly picking
the data plot from a lineup \(l\) is insufficient to determine the power
of a visual test because it does not provide information about the
number of selections made by the subject for the calculation of the
\(p\)-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
Therefore, we directly estimate the probability of a lineup being
rejected by assuming that individual skill has negligible effect on the
variation of the power. This assumption is not necessary true, but it
helps simplifying the model structure, thereby obviate a costly
large-scale experiment to estimate complex covariance matrices. The same
model given in Equation \ref{eq:logistic-regression-1-1} is applied to
model the power of a visual test.

>>>>

We can assess residuals in multiple ways. Residuals may be plotted as a
histogram or quantile-quantile plot to examine the distribution. Using
the classical normal linear regression model as an example, if the
distribution is symmetric and unimodal, we consider it to be well
behaved. However, if the distribution is skewed, bimodal, multimodal, or
contains outliers, there is a cause for concern. One could also inspect
the distribution by conducting a goodness-of-fit test, such as the
Shapiro-Wilk (SW) Normality test \citep{shapiro1965analysis}.

More typically, a scatterplot of residuals plotted against the fitted
values and each of the explanatory variables, is made to scrutinize
their relationships. If there are any visually discoverable patterns,
the model is potentially inadequate or incorrectly specifed. In general,
one looks for noticeable departures from the model like non-linear
dependency or heteroskedasticity. However, correctly judging a residual
plot where no pattern exists can be a difficult task for humans. It is
especially common, particularly among students, to misinterpret patterns
that are random noise and random deviation from a model
\citep{loy2021bringing}. There are also numerous conventional hypothesis
tests available to test for non-linear dependence
\citep{ramsey_tests_1969}, and heteroskedasticity
\citep{breusch_simple_1979}.

An abundance of literature describe appropriate diagnostic methods for
linear regression, e.g. \citet{draper1998applied},
\citet{montgomery1982introduction}, \citet{belsley_regression_1980},
\citet{cook_applied_1999} and \citet{cook1982residuals}. All these
writings consider plotting residuals to be an essential tool for
diagnosing regression modelling problems. In addition,
\citet{draper1998applied} and \citet{belsley_regression_1980} believe
that residual plots are usually revealing when the assumptions are
violated. \citet{cook_applied_1999} thinks formal tests and graphical
procedures are complementary and both have a place in residual analysis,
but they focus on graphical methods rather than on formal testing, as
they are easier to use. \citet{montgomery1982introduction} even suggests
that residual plots are more informative in most practical situations
than the corresponding conventional hypothesis tests.

@@

The power of a visual test on the other hand, may additionally depend on
the ability of the particular subject, as the skill of the individual
may affect the number of observers who identify the data plot from the
lineup \citep{majumder_validation_2013}. To address this issue,
\citet{majumder_validation_2013} models the probability of a subject
\(j\) correctly picking the data plot from a lineup \(l\) using a
mixed-effect logistic regression, with subjects treated as random
effect. Then, the estimated power of a visual test evaluated by a single
subject is the predicted value obtained from the mixed effects model.
However, this mixed effects model does not work with scenario where
subjects are asked to select one or more most different plots. In this
scenario, having the probability of a subject \(j\) correctly picking
the data plot from a lineup \(l\) is insufficient to determine the power
of a visual test because it does not provide information about the
number of selections made by the subject for the calculation of the
\(p\)-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
Therefore, we directly estimate the probability of a lineup being
rejected by assuming that individual skill has negligible effect on the
variation of the power. This assumption is not necessary true, but it
helps simplifying the model structure, thereby obviate a costly
large-scale experiment to estimate complex covariance matrices. The same
model given in Equation \ref{eq:logistic-regression-1-1} is applied to
model the power of a visual test.
